### 2020-09-15 Reports/JOURNAL - Confidence Regions/Conversa.tex

\documentclass[paper=a4, fontsize=11pt]{scrartcl}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amsthm,amssymb,bm,bbm}

\begin{document}

Empirical $p$-value of the sequence $\bm x'$ of length $N$ and embedding dimension $D$:
\begin{description}
    \item[Input: a sequence $\bm x'$ of length $N$ and  embedding dimension $D$]
    \item[Setup: $M$ reference points in the $H \times C$ plane produced by sequences of length $N$ and embedding dimension $D$]
    \begin{enumerate}
        \item Compute $(h',c')$ the point in the $H\times C$ plane produced by $\bm x'$
        \item Compute $b_{N,D}(\bm x')$: the smallest box which contains $(h',c')$ using the $M$ reference points
            \begin{itemize}
                \item the smallest $b_{N,D}(\bm x')$ is observed when $(h',c')$ ``coincides'' with the the point produced by the emblematic sequence;
                \item the largest box is any box associated to $(h',c')$ ``outside'' the largest box produced by the $M$ reference points
            \end{itemize}
        \item The empirical $p$-value of $\bm x'$ wrt the null hypothesis blablabla is the percentage of reference points outside $b_{N,D}(\bm x')$
    \end{enumerate}
\end{description}

Why one decision per configuration, and not a collective decision based on several values of $D$?

Imagine the situation of having a sample $\bm x$ and checking if the mean of the underlying distribution that produced the data is zero.
The simplest test statistic that we may use is a Student's-t Test.
But we could also use a test based on the median (the notches of the boxplot).
How do we integrate both evidences?
It is difficult!
It would lead us to Bonferroni-type tests, because we are using a single sample for several tests.

The focus of this paper is \textbf{not} assessing the quality of PRNG.
We are testing sequences in the configuration that appear in the literature.

The data either provide evidence to \textbf{reject} the null hypothesis, or not.
In the last case, we \textbf{do not} accept the null hypothesis; we just do not have enough evidence to reject it.

About the textures.
We assessed the trajectory of points $(h,c)$ computed over $z=\alpha w \cdot (1-\alpha) n_L$.
It will be interesting to see 
$z = m_\alpha (w) \cdot n_L$, where 
$m_\alpha$ is the function that maps $[0,1]\mapsto [\alpha,1]$.

\end{document}