### 2020-11-07 Reports/Report - Data Imputation/Code/EmpiricalDistributionTies.Rmd
Research Plan

---
title: "Study of the Empirical Distribution of Ties in Irrational Numbers"
output: html_notebook
---

The purpose of this study is describing the distribution of ties in irrational numbers when considering sequences of embedding dimension $D$ (and $\tau=1$).
The finding will help building a model for simulating such occurrences and, then, assessing imputation techniques in a Monte Carlo study.