### 2020-11-14 Reports/Report - Data Imputation/Code/EmpiricalDistributionTies.Rmd
Monte Carlo

---
title: "Study of the Empirical Distribution of Ties in Irrational Numbers"
author: "Eduarda Chagas"
date: "Nov 10, 2020"
output:
  html_document:
    df_print: paged
---

The purpose of this study is describing the distribution of ties in irrational numbers when considering sequences of embedding dimension $D$ (and $\tau=1$).
The finding will help building a model for simulating such occurrences and, then, assessing imputation techniques in a Monte Carlo study.
Thus, in this report we will make an exploratory analysis of the data, analyzing the percentage of tied sequences for each embedding dimension value.

### Load packages and sources

```{r} 
if(!require(ggpubr)){
  install.packages("ggpubr")
  require(ggpubr)
}
if(!require(ggplot2)){
  install.packages("ggplot2")
  require(ggplot2)
}
if(!require(ggthemes)){
  install.packages("ggthemes")
  require(ggthemes)
}
source('Bandt-Pompe.R')
```

### Reading the numbers

We have produced to files with Mathematica: e.txt, pi.txt, and sqrt2.txt. They contain expansions of these irrational numbers with $100,000$ digits.

```{r} 
e.data = as.character(read.table('../Data/e.txt', stringsAsFactors=FALSE, fileEncoding="latin1"))
pi.data = as.character(unlist(read.table('../Data/pi.txt', stringsAsFactors=FALSE, fileEncoding="latin1")))
sqrt2.data = as.character(unlist(read.table('../Data/sqrt2.txt', stringsAsFactors=FALSE, fileEncoding="latin1")))

e.vector = as.numeric(unlist(strsplit(e.data, ""))[3:100011])
pi.vector = as.numeric(unlist(strsplit(pi.data, ""))[3:100011])
sqrt2.vector = as.numeric(unlist(strsplit(sqrt2.data, ""))[3:100030])
```

### Computing the number of tied sequences for each value of $D\in\{3,4,5,6\}$

```{r} 
D = 3
Tau = 1

e.elements.D3 = formationPattern(e.vector, D, Tau, 1)
e.percent.D3 = percentual.equalities(e.elements.D3)

pi.elements.D3 = formationPattern(pi.vector, D, Tau, 1)
pi.percent.D3 = percentual.equalities(pi.elements.D3)

sqrt2.elements.D3 = formationPattern(sqrt2.vector, D, Tau, 1)
sqrt2.percent.D3 = percentual.equalities(sqrt2.elements.D3)

cat("Number of tied sequences \ne: ", round(e.percent.D3*100, 3), "%\npi: ", round(pi.percent.D3*100, 3), "%\nsqrt2: ", round(sqrt2.percent.D3*100, 3), "%\n")
```

```{r} 
D = 4
e.elements.D4 = formationPattern(e.vector, D, Tau, 1)
e.percent.D4 = percentual.equalities(e.elements.D4)

pi.elements.D4 = formationPattern(pi.vector, D, Tau, 1)
pi.percent.D4 = percentual.equalities(pi.elements.D4)

sqrt2.elements.D4 = formationPattern(sqrt2.vector, D, Tau, 1)
sqrt2.percent.D4 = percentual.equalities(sqrt2.elements.D4)

cat("Number of tied sequences \ne: ", round(e.percent.D4*100, 3), "%\npi: ", round(pi.percent.D4*100, 3), "%\nsqrt2: ", round(sqrt2.percent.D4*100, 3), "%\n")
```

```{r} 
D = 5
e.elements.D5 = formationPattern(e.vector, D, Tau, 1)
e.percent.D5 = percentual.equalities(e.elements.D5)

pi.elements.D5 = formationPattern(pi.vector, D, Tau, 1)
pi.percent.D5 = percentual.equalities(pi.elements.D5)

sqrt2.elements.D5 = formationPattern(sqrt2.vector, D, Tau, 1)
sqrt2.percent.D5 = percentual.equalities(sqrt2.elements.D5)

cat("Number of tied sequences \ne: ", round(e.percent.D5*100, 3), "%\npi: ", round(pi.percent.D5*100, 3), "%\nsqrt2: ", round(sqrt2.percent.D5*100, 3), "%\n")
```

```{r} 
D = 6
e.elements.D6 = formationPattern(e.vector, D, Tau, 1)
e.percent.D6 = percentual.equalities(e.elements.D6)

pi.elements.D6 = formationPattern(pi.vector, D, Tau, 1)
pi.percent.D6 = percentual.equalities(pi.elements.D6)

sqrt2.elements.D6 = formationPattern(sqrt2.vector, D, Tau, 1)
sqrt2.percent.D6 = percentual.equalities(sqrt2.elements.D6)

cat("Number of tied sequences \ne: ", round(e.percent.D6*100, 3), "%\npi: ", round(pi.percent.D6*100, 3), "%\nsqrt2: ", round(sqrt2.percent.D6*100, 3), "%\n")
```
### Analyzing the binary vectors with the position of the tied sequences

```{r} 
e.binary.D3 = binary.equalities(e.elements.D3)
e.binary.D4 = binary.equalities(e.elements.D4)
e.binary.D5 = binary.equalities(e.elements.D5)
e.binary.D6 = binary.equalities(e.elements.D6)

n.elements = 100

e.binary.df = data.frame('series' = c(e.binary.D3[1:n.elements], e.binary.D4[1:n.elements], e.binary.D5[1:n.elements], e.binary.D6[1:n.elements]),
                         'elements' = rep(c(1:n.elements), 4),
                         'D' = as.factor(c(rep(3, n.elements), rep(4, n.elements), rep(5, n.elements), rep(6, n.elements))))

ggplot(e.binary.df, mapping = aes(x = elements, y = series, group = D, color = D)) + 
    xlab("") + ylab("") +
    ggtitle("e number") +
    geom_line(position = position_dodge(0.8)) +
    theme_few(base_size = 13, base_family = "serif") + 
    facet_grid(facets = D~.) +
    scale_y_continuous(breaks=c(0, 1)) +
    theme(plot.title = element_text(hjust=0.5), legend.position = "none")
```

```{r} 
pi.binary.D3 = binary.equalities(pi.elements.D3)
pi.binary.D4 = binary.equalities(pi.elements.D4)
pi.binary.D5 = binary.equalities(pi.elements.D5)
pi.binary.D6 = binary.equalities(pi.elements.D6)

n.elements = 100

pi.binary.df = data.frame('series' = c(pi.binary.D3[1:n.elements], pi.binary.D4[1:n.elements], pi.binary.D5[1:n.elements], pi.binary.D6[1:n.elements]),
                         'elements' = rep(c(1:n.elements), 4),
                         'D' = as.factor(c(rep(3, n.elements), rep(4, n.elements), rep(5, n.elements), rep(6, n.elements))))

ggplot(pi.binary.df, mapping = aes(x = elements, y = series, group = D, color = D)) + 
    xlab("") + ylab("") +
    ggtitle("pi number") +
    geom_line(position = position_dodge(0.8)) +
    theme_few(base_size = 13, base_family = "serif") + 
    facet_grid(facets = D~.) +
    scale_y_continuous(breaks=c(0, 1)) +
    theme(plot.title = element_text(hjust=0.5), legend.position = "none")
```

```{r} 
sqrt2.binary.D3 = binary.equalities(sqrt2.elements.D3)
sqrt2.binary.D4 = binary.equalities(sqrt2.elements.D4)
sqrt2.binary.D5 = binary.equalities(sqrt2.elements.D5)
sqrt2.binary.D6 = binary.equalities(sqrt2.elements.D6)

n.elements = 100

sqrt2.binary.df = data.frame('series' = c(sqrt2.binary.D3[1:n.elements], sqrt2.binary.D4[1:n.elements], sqrt2.binary.D5[1:n.elements], sqrt2.binary.D6[1:n.elements]),
                         'elements' = rep(c(1:n.elements), 4),
                         'D' = as.factor(c(rep(3, n.elements), rep(4, n.elements), rep(5, n.elements), rep(6, n.elements))))

ggplot(sqrt2.binary.df, mapping = aes(x = elements, y = series, group = D, color = D)) + 
    xlab("") + ylab("") +
    ggtitle("sqrt2 number") +
    geom_line(position = position_dodge(0.8)) +
    theme_few(base_size = 13, base_family = "serif") + 
    facet_grid(facets = D~.) +
    scale_y_continuous(breaks=c(0, 1)) +
    theme(plot.title = element_text(hjust=0.5), legend.position = "none")
```

As we can see in the graphs above, the larger the dimension used, the greater the presence of patterns with repeated elements.
When we have $D = 3$, we see that there is a greater tendency for the existence of sequential patterns with the $label = 0$ (that is, without repeated elements).
However, as the dimension increases, this behavior is reversed.
This fact occurs because we are analyzing numbers within the small range $[0, 9]$, so the larger the symbol considered, the greater the probability of the existence of equal elements to be grouped.

### Examples of tied patterns

For illustration, follow we have presented some examples of tied sequences.
How we can observe, the implementation used for Bandt-Pompe symbolization considers the same approach of the time-ordered algorithm.

```{r} 
sqrt2.patterns.D3 = formationPattern(sqrt2.vector, 3, 1, 0)
sqrt2.binary.D3 = binary.equalities(sqrt2.elements.D3)
print('Elements:')
print(sqrt2.elements.D3[which(sqrt2.binary.D3 == 1)[1:10],])
print('Patterns:')
print(sqrt2.patterns.D3[which(sqrt2.binary.D3 == 1)[1:10],])
```
```{r} 
e.patterns.D4 = formationPattern(e.vector, 4, 1, 0)
e.binary.D4 = binary.equalities(e.elements.D4)
print('Elements:')
print(e.elements.D4[which(e.binary.D4 == 1)[1:10],])
print('Patterns:')
print(e.patterns.D4[which(e.binary.D4 == 1)[1:10],])
```
```{r} 
pi.patterns.D6 = formationPattern(pi.vector, 6, 1, 0)
pi.binary.D6 = binary.equalities(pi.elements.D6)
print('Elements:')
print(pi.elements.D6[which(pi.binary.D6 == 1)[1:10],])
print('Patterns:')
print(pi.patterns.D6[which(pi.binary.D6 == 1)[1:10],])
```

# Monte Carlo study

The purpose of the study is measuring the ability of imputation methods (list here which ones we will consider: I suggest the two best ones reported in the literature) to retrieve the underlying dynamic of a time series that has been *attacked*.
In a loose sense, we will assess the *breakdown point* of the imputation techniques (study references for "breakdown poiont").
The **attack** consists in introducing randomly repeated values in the sequence.

```{r PSEUDOCODE eval=FALSE}
# Define imputation techniques as functions
I1 <- function(time_series, params=...){return(imputed_time_series)}
I2 <- function(time_series){return(imputed_time_series)}

# Define attack
# Input:  "time_series" a time series without repetitions
#         "d" embedding dimension
#         "p" probability of attack
AttackTimeSeries <- function(time_series, d, p) {
  attacked_time_series <- time_series
  foreach(d in time_series) { # Check the package foreach
    if (runif(1) <= p) { # attack with probability p
      select uniformly an observation in d, say at position i
      select uniformly a position in d different from i, say j
      replace d[j] <- d[i] # this introduces a repetition
    }
  }
  return(attacked_time_series)
}

# Define the parameter space of the Monte Carlo study
# My suggestion: white noise, for which we already have confidence intervals in the HxC plane

N <- c(N1, N2, N3, ...) # lengths of the time series to be considered
D <- c(D1, D2, D3, ...) # embedding dimensions to be considered
P <- c(p1, p2, p3, ...) # probabilities of attack to be considered

# Store all the points in the HxC plane of the following loop
for(n in N){
  for(d in D) {
    for(p in P) {

      x <- white_noise_length_n # perhaps from the true noise we already have
      hcx <- point_in_the_HC_plane(x)
      x_attack <- AttackTimeSeries(x, d, p)
      hcxattack <- point_in_the_HC_plane(x_attack)

    }
  }
}

# Analyze the points, for instance:
# Measure distances between pairs
# Count the number of points inside the confidence regions
```