### 2021-10-10 Reports/Journal - Confidence Regions/Chapters/Conclusions.tex

\section{Conclusions}\label{Sec:Conclusions}

We presented and evaluated the first test for white noise in the  Entropy-Complexity plane.
Our proposal is based on two stages:
(1)~building nonparametric empirical confidence regions in the principal components space and mapping these boxes back to the $H\times C$ plane.
(2)~computing an approximate $p$-value for a given sample by comparing it with points produced by true white noise random sequences (TWNRS).
We obtained the TWNRS with data from physical devices.

Our test has a good size, mostly with long TWNRS.
We also determined the power of our test for the alternative hypothesis of correlated $f^{-k}$ noise and found that it rejects the null hypothesis ($k=0$) for $k>3/10$.

Although our work focuses on the study of short sequences, we were able to capture the random behavior of well-known pseudorandom number generators already analyzed in the literature. 
With this, we verified the adequacy of our technique as it is capable of detecting correlation structures.